---
title: "Ad Hoc Wireless Network"
author: "Elvin Liu (2239429)"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
```

```{r genNodes}
set.seed(123)

# Generate nodes in a given two-dimensional ad hoc network
# (n is the number of nodes to generate)
genNodes = function(n) {
  # nodeDensity() returns a numeric vector of values that are
  # proportional to the node density value at the (x, y) pair
  source("nodeDensity.R")
  
  # Repeatedly take the node density to find a
  # value close to the absolute largest possible
  nodeDensity_max = max(replicate(1000,
                                  nodeDensity(runif(1, 0, 100),
                                              runif(1, 0, 100))))
  
  mat = matrix(NA, nrow = n, ncol = 2)
  
  # Use acceptance-rejection sampling to randomly generate the nodes of the
  # network according to the pre-specified node density function nodeDensity()
  while (n > 0) {
    x = runif(1, 0, 100)
    y = runif(1, 0, 100)
    u = runif(1, 0, nodeDensity_max)
    
    if (u < nodeDensity(x, y)) {
      mat[n, 1] = x
      mat[n, 2] = y
      n = n - 1
    }
  }
  
  # return value is an n by 2 matrix with the first column representing the
  # x-coordinates and the second column as the y-coordinates of the nodes
  return(mat)
}
```

```{r findTranMat()}
# Finds the transition matrix based on a distance matrix mat
# (stores all pairwise distances in the network) and the power level \bar{R}
findTranMat = function(mat, R) {
  n <- nrow(mat)
  tranMat <- matrix(0, nrow = n, ncol = n)
  
  for (i in 1:n) {
    
    # Any pair of nodes whose distance is <= \bar{R} is connected
    Si <- which(mat[i,] <= R)
    num_connected <- length(Si)
    
    if (num_connected > 0) {
      tranMat[i, Si] <- 1 / num_connected
    }
  }
  
  return(tranMat)
}
```

```{r getEigen2()}
# Size of the second largest eigenvalue is the key
# to determine if network is connected or not

# Helper function getEigen2() finds it
# Input is an n by n transition matrix
getEigen2 = function(mat) {
  e_values <- eigen(mat)$values
  
  eigen2 <- e_values[2]
  
  # returns a numeric value
  # magnitude of the second largest eigenvalue of the input matrix
  return(eigen2)
}
```

```{r findRange()}
# Keep in mind diagonals are 0

# Also note the smallest possible value that R_c can take is:
#   R_c >= max{a_1, ..., a_n} with a_i = min(D_{i1}, ..., D_{in})

# The largest possible value that R_c can take is:
#   R_c <= min{b_1, ..., b_n} with b_i = max(D_{i1}, ..., D_{in})

# Helper function which finds the range of possible values to search for R_c
# based on the above observations...
#   a) There is only one input argument of this function as max,
#   which is the distance matrix D
#   b) The function returns a numeric function of length 2,
#   which are the minimum and maximum values to search for R_c
findRange = function(mat) {
  
  # Set minimum to as large as possible and maximum to as small as possible
  rowMin = rep(Inf, nrow(mat))
  rowMax = rep(-1 * Inf, nrow(mat))
  
  # Find row minimums and maximums
  for (r in 1:nrow(mat)) {
    if (r == 1) {
      rowMin[r] = mat[r, 2]
      rowMax[r] = mat[r, 2]
    }
    else {
      rowMin[r] = mat[r, 1]
      rowMax[r] = mat[r, 1]
    }
    
    # Set minimums and maximums according to the note
    for (c in 1:nrow(mat)) {
      if (c != r) {
        if (rowMin[r] > mat[r, c]) {
          rowMin[r] = mat[r, c]
        }
        if (rowMax[r] < mat[r, c]) {
          rowMax[r] = mat[r, c]
        }
      }
    }
  }
  
  # find the smallest \bar{R} within a certain
  # tolerance level that gives a connected network
  c(max(rowMin), min(rowMax))
}
```

```{r findRc}
# For a given configuration of nodes, we intend to find the
# smallest radius R_c > 0 such that the network is connected

# (i.e. there is a path passing through nodes
# in the network between each pair of nodes)

# Outputs R_c for a given collection of nodes

# nodes is a two-column matrix of the x and y coordinates/
# locations of the nodes (e.g. generated from the function genNodes())

# tol (default value 0.05) is the tolerance level for how close we need
# to get to the true value of R_c for the provided node configuration
findRc = function(nodes, tol = 0.05) {
  distance_mat <- as.matrix(dist(nodes))
  
  range_r <- findRange(distance_mat)
  
  # Bisection method
  r_min <- range_r[1]
  r_max <- range_r[2]
  r_mid <- (r_min + r_max) / 2
  
  while (abs(r_min - r_max) >= tol) {
    r_mid = (r_min + r_max) / 2
    
    tranMat <- findTranMat(distance_mat, r_mid)
    
    eigen2 <- Mod(getEigen2(tranMat))
    
    # if midpoint gives connected network, assign r_max as midpoint
    if(abs(eigen2 - 1) > 1e-10 & eigen2 < 1) {
      r_max = r_mid
    }
    else {
      # otherwise, set r_min as midpoint
      r_min = r_mid
    }
  }

  # numeric variable that holds the value of R_c,
  # or a value close to it under the tolerance level
  return(r_mid)
}
```

```{r Testing genNodes()}
# Generate n = 10000 nodes from genNodes() and use
# a scatter plot to visualize the generated nodes

# This visualization helps you confirm that the function genNodes() is indeed
# generating nodes according to the required density function nodeDensity()

a = genNodes(10000)
plot(a[,1], a[,2], main = "Scatter Plot of Node Densities for n = 10000", xlab = "x-position", ylab = "y-position", pch = 20)
```

```{r}
# Select a range of sample sizes for n
# For each n, generate 1000 networks via genNodes()
n <- c(50, 100, 150, 200, 250)
N <- 1000

# Find the value of R_c for each of these 1000 networks
hist <- lapply(n, function(x) {
  replicate(N, findRc(genNodes(x), tol = 0.001))
})

combined_data <- bind_rows(
  lapply(seq_along(hist),
         function(i) {
           data.frame(value = hist[[i]],
                      source_list = paste("n = ", i * 50))
}))

# Examine the distribution of these R_c values for each n
ggplot(combined_data, aes(x = value, color = source_list)) +
  geom_density() + geom_rug() +
  labs(x = expression(R[c]),
       y = "Densities",
       title = paste("Distribution of", expression(R[c]),
                     "for Various Node Sample Sizes"),
       color = "Sample Size (n)")
```

(a) How does $R_c$, the smallest radius such that the network is connected, change with different node configurations?

> At larger inputs of $n$ in genNodes(), the smallest radius such that the network is connected ($R_c$) is, on average, smaller.

(b) Explore the distribution of $R_c$ (for each $n$). Is it symmetric, skewed, heavy-tailed, and multimodal.

> For small values of $n$ in genNodes(), the distribution of $R_c$ is pretty symmetric, and grows more and more right-skewing as $n$ increases. It is not particularly heavy-tailed. They are all generally unimodal with only some small bumps in the graphs.

```{r}
# For n = 50, plot the network of connected nodes as well as their associated
# edges for four of your 1000 n-node configurations corresponding roughly to
# the minimum, mean, median, and maximum values of R_c
size <- 50
N <- 1000

set.seed(123)
seed_list <- sample(1:100000, N, replace = TRUE)

mat1 <- matrix(data = 0, ncol = 2, nrow = N)

for (i in 1:length(seed_list)) {
  curr_seed = seed_list[i]
  set.seed(curr_seed)
  Rc = findRc(genNodes(size), tol = 0.001)
  mat1[i, ] = c(curr_seed, Rc)
}

# Median
median_single <- function(x) {
  if (length(x) %% 2 == 0) {
    sorted_x <- sort(x)
    return(sorted_x[(length(x) + 1) / 2])
  } 
  else {
    return(median(x))
  }
}

# Mean
closest_mean <- function(x) {
  sorted_x <- sort(x)
  mean_x <- mean(x)
  
  closest_x <- x[1]
  closest_diff <- abs(closest_x - mean_x)
  
  for(i in 2:length(x)) {
    curr_x = x[i]
    curr_diff = abs(closest_x - mean_x)
    if (curr_diff < closest_diff) {
      closest_x = curr_x
      curr_diff = curr_diff
    }
  }
  
  return(closest_x)
}

# Matrices for minimum, median, maximum, and mean respectively
s1 <- mat1[min(mat1[, 2]) == mat1[, 2]][1]
s2 <- mat1[median_single(mat1[, 2]) == mat1[, 2]][1]
s3 <- mat1[max(mat1[, 2]) == mat1[, 2]][1]
s4 <- mat1[closest_mean(mat1[, 2]) == mat1[, 2]][1]
```

```{r}
connectedGraph = function(s, size = 50, name) {
  set.seed(s)
  
  nodes <- genNodes(size)
  
  Rc <- findRc(nodes, tol = 0.0001) + 0.005
  
  # dist() is euclidean distance between nodes
  distances <- as.matrix(dist(nodes))
  
  # Most edges in a simply connected graph is n * (n - 1) / 2 for n vertices
  pairs <- rbind(cbind(nodes,
                        matrix(data = NA,
                               nrow = nrow(distances),
                               ncol = 2)),
                 matrix(data = NA, nrow = (size^2) / 2, ncol = 4))
  
  # extra_pairs keeps track of the index of the pairs past the initial 50
  extra_pairs <- size + 1
  for (i in 1:nrow(distances)) {
    for (j in 1:ncol(distances)) {
      if (j > i) {
        if (distances[i, j] <= Rc) {
          if (is.na(pairs[i, 3])) {
            pairs[i, 3:4] = nodes[j, ]
          }
          else {
            pairs[extra_pairs, 1:2] = nodes[i, ]
            pairs[extra_pairs, 3:4] = nodes[j, ]
            extra_pairs = extra_pairs + 1
          }
        }
      }
    }
  }
  
  # x and y positions of the nodes
  df <- data.frame(x = nodes[, 1], y = nodes[, 2])

  # pairs of nodes that are connected by an edge
  end <- data.frame(x = pairs[, 1], y = pairs[, 2],
                    x_end = pairs[, 3], y_end = pairs[, 4])
  
  end <- na.omit(end)
  
  graph <- ggplot(data = df, mapping = aes(x = x, y = y)) +
    geom_point() +
    geom_segment(data = end,
                  mapping = aes(x = x, y = y,
                                xend = x_end, yend = y_end, color = "")) +
    scale_color_manual(values = "green") +
    theme_bw() +
    labs(x = "x-coordinates of nodes",
         y = "y coordinates of nodes",
         title = paste("Node Configurations of Size 50 with", name,
                       expression(R[c]), "=", round(Rc, 4)))
  
  return(graph)
}

# Connected graphs for Minimum, Median, Maximum, and Mean respectively
connectedGraph(s = s1, name = "Minimum")
connectedGraph(s = s2, name = "Median")
connectedGraph(s = s3, name = "Maximum")
connectedGraph(s = s4, name = "Mean")
```

```{r Scratch Work}
# Naive way that only works for s1:

# set.seed(s1)
# s1_nodes <- genNodes(size)
# s1_Rc <- findRc(s1_nodes, tol = 0.001)
# s1_distances <- as.matrix(dist(s1_nodes))
# 
# s1_pairs <- rbind(cbind(s1_nodes, matrix(data = NA, nrow = nrow(s1_distances), ncol = 2)), matrix(data = NA, nrow = size^2/2, ncol = 4))
# 
# extra_pairs <- size + 1
# for (i in 1:nrow(s1_distances)) {
#   for (j in 1:ncol(s1_distances)) {
#     if (j > i) {
#       if (s1_distances[i, j] <= s1_Rc) {
#         if (is.na(s1_pairs[i, 3])) {
#           s1_pairs[i, 3:4] = s1_nodes[j, ]
#         }
#         else {
#           s1_pairs[extra_pairs, 1:2] = s1_nodes[i, ]
#           s1_pairs[extra_pairs, 3:4] = s1_nodes[j, ]
#           extra_pairs = extra_pairs + 1
#         }
#       }
#     }
#   }
# }
# 
# s1_df <- data.frame(x = s1_nodes[, 1], y = s1_nodes[, 2])
# 
# s1_end <- data.frame(x = s1_pairs[, 1], y = s1_pairs[, 2], x_end = s1_pairs[, 3], y_end = s1_pairs[, 4])
# s1_end <- na.omit(s1_end)
# 
# ggplot(data = s1_df,
#        mapping = aes(x = x, y = y)) +
#   geom_point() +
#   geom_segment(data = s1_end,
#                mapping = aes(x = x, y = y, xend = x_end, yend = y_end, color = "red"))

# Method using adjacency matrix:

# # Distance between nodes
# compute_distance <- function(node1, node2) {
#   return(sqrt((node1[1] - node2[1])^2 +
#               (node1[2] - node2[2])^2))
# }
# 
# num_nodes <- nrow(s1_nodes)
# adj_matrix <- matrix(0, nrow = num_nodes, ncol = num_nodes)
# 
# # Populate adjacency matrix based on distance criterion
# for (i in 1:num_nodes) {
#   for (j in 1:num_nodes) {
#     if (j >= i) {
#       if (compute_distance(s1_nodes[i, ], s1_nodes[j, ]) <= s1_Rc) {
#         adj_matrix[i, j] <- 1
#         adj_matrix[j, i] <- 1
#       }
#     }
#   }
# }
# 
# graph <- graph_from_adjacency_matrix(adj_matrix)
```



